\documentclass[letterpaper]{article}

\usepackage{geometry}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{titlesec}


\setlength\parindent{0em}


\begin{document}

\begin{enumerate}

\section*{Equations and Inequalities}

\item Express $f(x) = |x-1| + |x+1|$ as a piecewise function. Sketch the graph of $y=f(x)$. Hence, or otherwise, solve the inequality

$$
|x-1| + |x+1| \ge x+2
$$

\item The graph of $y=h(x)$ intersects the graph of $y=h^{-1}(x)$ at $x=2$ and $x=3$. Given that $h(x) = x^3 + ax + b$, find the values of $a$ and $b$.

\item Find the value of $p$ if the roots of the quadratic equation $4x^2 - x + 5p = 5px$ differ by 1, and $p$ is a constant real number.

\item Given that the roots of the quadratic equation $2x^2 - x - 4$ are $\alpha$ and $\beta$, find a quadratic equation whose roots are $3\alpha + \beta$ and $\alpha + 3\beta$.

\item Show that $x=1$ is a solution of $x^3 - 1 = 0$. Hence or otherwise, factorize $x^3 - 1$.

\item Find the term independent of $x$ in $\left(x + \cfrac{1}{x}\right)^{2n}$ where $n$ is an integer.

\item Show, without a calculator, that $2 < \sqrt{5} < 3$, and deduce that $0 < \sqrt{5} - 2 < 1$.

Show that $(\sqrt 5 + 2)^4 (\sqrt 5 - 2)^4 = 332$. 

Hence or otherwise, show that $321 < (\sqrt 5  +2)^2 < 322$.

\item Express $\cfrac{x^2 - 1}{x^2 - 16}$ as a partial fraction.

\item Solve the simultaneous equation

\begin{align*}
\log_5 &p + \log_5 q = 3 \\
\log_p &q = 2
\end{align*}

\item Solve the simultaneous equation

\begin{align}
16^x - 16^y &= 64512 \\
4^x - 4^y &= 224
\end{align}

\item Given that the roots of $(\log 3x)(\log 7x) = 1$ are $\alpha$ and $\beta$, find $\alpha\beta$.

\section*{Trigonometry}

\item Find the exact values of 

\begin{enumerate}
\item $\sin{\cfrac{\pi}{12}}, \cos{\cfrac{\pi}{12}}$ and $\tan{\cfrac{\pi}{12}}$
\item $\cos^4{\cfrac{\pi}{24}} - \cos^4{\cfrac{\pi}{24}}$
\item $\tan^{-1} 1 + \tan^{-1} 2 + \tan^{-1} 3$
\end{enumerate}

\item Prove the following identities.

\begin{enumerate}
\item $\tan \theta + \cot \theta = 2\mathrm{cosec}\ 2\theta$
\item $\tan 2A - 2 \tan 2A \sin^2 A = \sin 2A$
\end{enumerate}

\section*{Calculus}

\item The equation of a curve is $y = e^{-ax^2}$, where $a > 0$ is a constant real number. 

\begin{enumerate}
\item Find $\cfrac{dy}{dx}$ and $\cfrac{d^2y}{dx^2}$
\item Find the stationary points of the curve. Over which interval is the function strictly increasing?
\end{enumerate}

\item In an experiment, synthetic crude oil is piped into a silo at a rate of 

$$
R(t) = \frac{70}{\sqrt{t}}
$$

tonnes per second, from $t=1$ onwards. However, oil is leaking from the silo at a rate of

$$
C(t) = \sqrt{t} + 9
$$

tonnes per second. At $t=1$ second, the amount of oil in the silo is 10,000 tonnes.

\begin{enumerate}
\item Find the net change in the amount of oil in the silo between $t=4$ and $t=16$ seconds.
\item Write down an expression involving an integral that gives the amount of synthetic crude oil inside the silo at time $t=20$ second.
\end{enumerate}

\item A cylinder has radius $\frac{t^{3/2}}{1+t^{3/2}}$ and height $\frac{t}{1+t}$, where $t$ is the time in seconds. What is the rate of change of its volume? What is the rate of change of its surface area (including top and base)?

\item A chemist has a filter paper in the shape of a disc of radius $R$. The chemist cuts out a slice of the disc that subtends an angle $\theta$ and folds the remaining piece into a cone. Find the volume of the cone as a function of $\theta$ and $R$. What is the maximum volume of the cone? For what value of $\theta$ does the cone attain this volume?

\end{enumerate}
\end{document}
