\documentclass[letterpaper]{article}

\usepackage{geometry}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{pgfplots}


\setlength\parindent{0em}


\begin{document}

\section{Introduction}

Definition of relation: association between elements of two sets. Arrow diagram. Four types of relations - one/many to one/many. Definition of domain/range. Definition of function. Dependent and independent variables. ``Mapping''. Representation of a function - description, tabulation, graph, arrow diagram, formula. Notation for defining a function.

\begin{enumerate}
\item A function is defined by $g: x \to 3x^2 + 5$. Find the images of $-1$ and $4$. Find also $g(3a)$ and $g(2p-1)$.

\item A function $f: x \to \cfrac{72}{ax + b}, x \ne -\cfrac{b}{a}$, maps 7 to 9 and 6 to 12.
\begin{enumerate}
\item Explain why the definition of $f$ specifies $x \ne -cfrac{b}{a}$.
\item Find the value of $a$ and $b$.
\item Find the element that under this mapping has an image of 4.
\end{enumerate}

Vertical line test. Domain and Range. Definition of Invariant. Look at graph to find the range!

\item For each of the functions below, state if there is any restriction on the domain, and write down the domain in algebraic and interval form.

\begin{enumerate}
\item $f(x) = 3x^2 - 5$
\item $g(x) = \sqrt{x-2}$
\item $h(x) = \cfrac{x}{3x + 1}$
\end{enumerate}

\item Find the range of each of the following functions. Find also the value(s) of $x$ which is invariant under each of the functions.

\begin{enumerate}
\item $f: x \to 2x^2 + 1, x \in R$
\item $g: x \to 2x^2 + 1, -1 \le x \le 3$
\item $h: x \to 2x^2 + 1, x \in R^+$
\end{enumerate}

Function is not defined unless the domain is specified. If unspecified, take to be largest possible domain. What is the larges tpossible domains for $f, g, h$ above?

\section{Families of Functions}

Even and odd function. 

Determine algebraically if the functions below are even, odd or neither. Use your GC to graph the functions to check.

\begin{enumerate}
\item $f(x) = 3x^2 - 5$
\item $g(x) = 2x+ x^3$
\item $h(x) = \cfrac{x}{3x + 1}$
\end{enumerate}

For each of the following functions, determine if it is odd, even, or neither. Specify the interval over which the function is increasing, decreasing or constant. State where an extreme value occurs. State the range of the function.

\begin{enumerate}
\item $f: x \to x$
\item $f: x \to 3 - 2x$
\item $f: x \to 2x^2 - 4$
\item $g: x \to x(3-x)$
\item $f: x \to 2 - 3x^3$
\item $g: x \to (x-1)(x+2)(x+3)$
\item $f(x) = \sqrt{x}$
\item $g(x) = \cfrac{1}{x}$
\item $h(x) = 2^x$
\item $f(x) = \log x$
\end{enumerate}

Given $f(x) = |2x + 7|$, evaluate $f(-5)$ and $f(\cfrac{1}{4})$, and find the possible values of $x$ given that $f(x) = 6$.

Sketch the given functions and state the range. $f: x \to |3x - 5|, -3 \le x < 4$ and $g: x \to |2 - x|$.

Sketch the function $f(x)$ and state the domain, range and intervals over which the fnction is increasing, decreasing or constant, where $f(x) = x^2 - 2, -3 \le x < 0, 3x - 2, 0 \le x < 2, 2^x, 2 \le x < 4$. Also for $g(x) = |x|, -5 \le x < 1, \ln x, x \ge 1$

A curve is defined parametrically by the equations $x = 2 - t^2, y = t(t+1)$. Find the points $P$ and $Q$ on the curve with parameters $-2$ and $3$. Find the length of $PQ$.

For each of the parametric equations given below, find the Cartesian equations. Determine whether the equation of the curve is a function. If it is, giv the maximum domain and range of the function.

\begin{enumerate}
\item $x = 4t, y = 3t^2$
\item $x = t^2 - 1, y = t+1$
\item $x = 2+t, y = \cfrac{3+t}{t+1}, t \ne -1$
\end{enumerate}

Which of the following are 1-1 functions?

\begin{enumerate}
\item $f: x \to \sqrt{x}, x \ge 0$
\item $g: x \to \cfrac{1}{x}, x \ne 0$
\item $h: x \to |x|$

For the functions which are not one-to-one, give a suitable domain for he function so that it becomes an injective function.

A function $f$ is given by $f: x \to \cfrac{x - 5}{2x - 4}, x \ne 2$. Find the values of $f^{-1}(1)$ and $f^{-1}(3)$. Find the value of $a$ for which $f^{-1}(a) = 2a + 1$.

Find $f^{-1}(x)$ for the function $f: x \to \cfrac{x-3}{x+1}, x \ne -1$.

Find the inverse function for

\begin{enumerate}
\item $f(x) = 2x^2 + 4, x > 0$
\item $g(x) = \cfrac{2x + 5}{x - 7}, x \ne 7$
\item $h(x) = x^2 + 2x, x \ge -1$
\item $m(x) = x + \cfrac{1}{x}, x \ge 1$
\end{enumerate}

Given $f: x \to \cfrac{1}{1 + 3x}, x > -\cfrac{1}{3}$, find

\begin{enumerate}
\item The range of $f$
\item The inverse function $f^{-1}$, stating the corresponding range
\item Sketch the graphs of $f$ and $f^{-1}$ on the same diagram.
\end{enumerate}

Let $f: R \to R$ be defined by $f: x \to x^2 - 1, x \gt 0, 3x + 4, x < 0$. Sketch the graph of $f$. Show that $f^{-1}$ exists and define it. Hence sketch the graph of $f^{-1}$ and find the value of $x$ for which $f(x) = f^{-1}(x)$.

Functions $f$ and $g$ are defined by $f(x) = px - 2$ where $p$ is a constant and $g(x) = 4x + 3$. Find an expression for $f \circ g$ and the value of $p$ for which $f \circ g = g \circ f$.

For each of the following functions $f$ and $g$, obtain an expression for $f \circ g$ and $g \circ f$. Consider if there is any restriction on the domain.

\begin{enumerate}
\item $f: x \to 2x + 1, g: x \to 3 + x^2$
\item $f: x \to 4 - x, g: x \to \cfrac{2x}{x - 3}, x \ne 3$
\item $f: x \to x^3, g: x \to x + 2$
\end{enumerate}



\end{enumerate}


\end{enumerate}

\end{document}
